% Appendix A2.1 / A2.2 — full proofs of Proposition 3 and Lemma 3.
% Source-of-truth: project/plans/Prop3_Lemma3_visienhance_theory.md (audited 2026-05-27,
% canonical form: √ε scaling, β = M L_q / √2). Numerical toy: tests/test_theorems_numerical.py.

\section{Proofs for Diagnosis-Preserving Enhancement}\label{app:a2}

\subsection{Assumptions}\label{app:a2-assump}
\begin{assumption}[Quality improvement]\label{as:a1}
$\Tw$ satisfies $\qbar(\Tw(x,q)) > \qbar(x)$ almost surely on the salvage band
$\qbar\in[\tauenh,\tauhigh]$ (outside this band no improvement is guaranteed).
\end{assumption}
\begin{assumption}[Calibrated quality scoring]\label{as:a2}
The frozen scorer $f_\eta:\mathcal{X}\to[0,1]^5$ yields $\qbar$ monotone in ground-truth
degradation severity on paired data (PLCC $0.924$, SRCC $0.895$).
\end{assumption}
\begin{assumption}[Monotone prior]\label{as:a3}
The \qvib{} prior variance $\sigma^2(\qbar)$ is strictly decreasing in $\qbar$ on $(0,1)$
(Lemma~1), i.e.\ $\tfrac{d\sigma^2}{d\qbar} < 0$.
\end{assumption}
\begin{assumption}[Diagnostic preservation]\label{as:a4}
$\Tw$ is trained with a DP loss $\Ldp\leq\epsilon$ (Lemma~\ref{lem:dp-full}), so
$p_\phi(z\mid x_{\mathrm{enh}},q')$ and $p_\phi(z\mid x_{\mathrm{ref}},q_{\mathrm{ref}})$
are $\epsilon$-close in KL.
\end{assumption}

% ═══════════════════════════════════════════════════════════════════════════════
\subsection{Proposition 3 (Variance-Driven Entropy Term, Sketch)}\label{app:a2-prop3}

\paragraph{Scope and empirical status.}
We sketch the directional effect of the quality-adaptive prior variance on predictive
entropy under enhancement. We emphasise at the outset that the \emph{net} entropy of the
predictive distribution is \textbf{not} empirically supported to decrease under
enhancement: a quality-conditioned predictive-entropy probe finds that the mean entropy
\emph{increases} after enhancement on the mixed-severity test split
($\bar{\Hent}_{\mathrm{deg}}=0.1487\!\to\!\bar{\Hent}_{\mathrm{enh}}=0.1572$; two
independent probes agree). The derivation below therefore isolates only the
$\sigma^2$-driven term $-\tfrac{d}{2}\log\tfrac{\sigma^2(\qbar)}{\sigma^2(\qbar')}$ and
its direction; it does \emph{not} establish a net entropy reduction, because the
residual and drift terms are not controlled tightly enough to dominate empirically.
Consequently, the role that Proposition~3 plays in the paper --- linking enhancement to
preserved discriminative information --- is carried not by a net-entropy claim but by
(i)~Lemma~\ref{lem:dp-full}, whose mutual-information lower bound is empirically validated
(experiment E7: $\Delta\mathrm{AUC}_{\mathrm{enh}}=+0.0205$, $95\%$ CI $[+0.005,+0.035]$,
McNemar $p=2.3\times10^{-45}$), and (ii)~the Stage-1 PSNR$\,\geq 30$\,dB non-emptiness
condition on the salvage band (Assumption~\ref{as:a1}). See the Limitations discussion in
the main text, which states that these arguments are sketches rather than fully watertight
proofs.

\begin{proposition}[$\sigma^2$-driven entropy term, directional]\label{prop:entropy-full}\label{prop:entropy}
Under Assumptions~\ref{as:a1}--\ref{as:a4}, for $x$ drawn from the salvage band and
$q'=f_\eta(\Tw(x,q))$, the variance-driven component of the expected predictive entropy
satisfies the directional bound
\begin{equation}
\E_{x\sim p(x\mid\qbar)}\!\big[\Hent(\hat{p}_{\Tw(x,q)})\big] \;\leq\;
\E_{x\sim p(x\mid\qbar)}\!\big[\Hent(\hat{p}_x)\big]
- \tfrac{d}{2}\log\tfrac{\sigma^2(\qbar)}{\sigma^2(\qbar')} + 2\beta\sqrt{\epsilon},
\label{eq:p3-full}
\end{equation}
\emph{provided} the unmodelled mean-drift contribution $\Delta\Hent_\mu$ (Step~4) is
non-positive in expectation. We stress that this proviso is \textbf{not} met empirically
(see the scope note above), so \eqref{eq:p3-full} should be read as characterising the
\emph{sign} of the $\sigma^2$-driven term, not as a certificate of net entropy reduction.
\end{proposition}

\begin{proof}[Derivation (sketch, under Assumptions~\ref{as:a1}--\ref{as:a4}; see the
main-text Limitations for non-watertight regimes)]
We sketch the argument in five steps; it isolates the direction of the
$\sigma^2$-driven term rather than proving a net reduction.

\paragraph{Step 1 (ELBO decomposition of predictive entropy).}
By Jensen's inequality and the \qvib{} ELBO (Proposition~1),
\begin{equation}
\Hent(\hat{p}_x) = \Hent\!\big(\E_{z\sim p_\phi(z\mid x,q)}[q_\theta(y\mid z)]\big)
\;\geq\; \E_{z\sim p_\phi}\!\big[\Hent(q_\theta(y\mid z))\big] - I(Z;Y\mid X=x).
\label{eq:p3-1}
\end{equation}
The Markov chain $X\to Z\to Y$ keeps $I(Z;Y\mid X=x)$ small, so the dominant term is
$\E_z[\Hent(q_\theta(y\mid z))]$.

\paragraph{Step 2 (Bound conditional entropy by encoder variance).}
A second-order Taylor expansion of $\Hent(q_\theta(y\mid z))$ around the posterior mean
$\mu_\phi(x,q)$ gives, with $\Sigma_\phi=\mathrm{diag}(\sigma_\phi^2(x,q))$,
\begin{equation}
\E_z[\Hent(q_\theta(y\mid z))] = \Hent(q_\theta(y\mid\mu_\phi))
+ \tfrac{1}{2}\sum_{j=1}^{d}\sigma_{\phi,j}^2\,\partial^2_{z_j}\Hent + O(\|\sigma_\phi\|^4).
\label{eq:p3-2}
\end{equation}

\paragraph{Step 3 (KL forces $\sigma_\phi^2$ to track $\sigma^2(\qbar)$).}
The \qvib{} KL regulariser is
$\KL(p_\phi(z\mid x,q)\,\|\,r_\psi(z\mid q))
= \tfrac{1}{2}\sum_j[\tfrac{\mu_{\phi,j}^2+\sigma_{\phi,j}^2}{\sigma^2(\qbar)}
- 1 - \log\tfrac{\sigma_{\phi,j}^2}{\sigma^2(\qbar)}]$.
At training equilibrium the first-order condition in $\sigma_{\phi,j}^2$ yields
$\sigma_{\phi,j}^2 = \sigma^2(\qbar)\,(1+\mathcal{O}(\beta^{-1}))$. Substituting into
\eqref{eq:p3-2},
\begin{equation}
\E_z[\Hent(q_\theta(y\mid z))] \approx \Hent(q_\theta(y\mid\mu_\phi))
+ \tfrac{d}{2}\sigma^2(\qbar)\,c(x,q) + O(\sigma^4),
\label{eq:p3-3}
\end{equation}
with $c(x,q)=\tfrac{1}{d}\sum_j\partial^2_{z_j}\Hent|_{z=\mu_\phi}\geq 0$ on average.

\paragraph{Step 4 (Compare $\qbar$ versus $\qbar'$).}
For the enhanced input $x_{\mathrm{enh}}=\Tw(x,q)$ with $\qbar'>\qbar$ (Assumptions~\ref{as:a1},
\ref{as:a2}),
\begin{equation}
\E_z[\Hent(q_\theta(y\mid z_{\mathrm{enh}}))] - \E_z[\Hent(q_\theta(y\mid z_x))]
\approx \tfrac{d}{2}\big(\sigma^2(\qbar') - \sigma^2(\qbar)\big)\,\bar{c} + \Delta\Hent_\mu,
\label{eq:p3-4}
\end{equation}
where $\bar{c}\geq 0$ and $\sigma^2(\qbar')<\sigma^2(\qbar)$ by Assumption~\ref{as:a3} makes the
first term strictly negative. The mean drift $\Delta\Hent_\mu$ is \emph{bounded in magnitude}
by Assumption~\ref{as:a4}: Lemma~\ref{lem:dp-full} gives
$\|p_\phi(z\mid x_{\mathrm{enh}},q') - p_\phi(z\mid x_{\mathrm{ref}},q_{\mathrm{ref}})\|_{\TV}
\leq \sqrt{\epsilon/2}$, and Fannes-style continuity with Lipschitz $q_\theta$ yields
$\|\Delta\Hent_\mu\| \leq 2\beta\sqrt{\epsilon}$ (same Pinsker constant as Lemma~\ref{lem:dp-full}).
We caution that this controls only the \emph{magnitude} of $\Delta\Hent_\mu$, not its sign:
the bound permits $\Delta\Hent_\mu$ to be positive, in which case the mean-drift contribution
can offset or exceed the negative variance-driven term. This is precisely the regime our
measurements indicate (net entropy increases under enhancement), so we do not claim that the
right-hand side of \eqref{eq:p3-4} is negative in practice.

\paragraph{Step 5 (Refine via Lemma~1 monotonicity).}
With $\sigma^2(\qbar)=\sigma_{\min}^2+(\sigma_{\max}^2-\sigma_{\min}^2)(1-\sigmoid(s(\qbar-\qbar_0)))$,
for $\qbar'=\qbar+\Delta\qbar$ with $\Delta\qbar\in[0.1,0.25]$,
\begin{equation}
\log\tfrac{\sigma^2(\qbar)}{\sigma^2(\qbar')}
\geq s\,\Delta\qbar\,(1-\sigmoid(\cdot))\,\tfrac{\sigma_{\max}^2-\sigma_{\min}^2}{\sigma_{\max}^2}
= \Theta(s\,\Delta\qbar) > 0.
\label{eq:p3-5}
\end{equation}
Combining \eqref{eq:p3-4}--\eqref{eq:p3-5} and taking the expectation over $x$ in the
salvage band gives the directional statement \eqref{eq:p3-full}, under the proviso (stated in
the proposition) that $\Delta\Hent_\mu$ is non-positive in expectation. Since that proviso is
not met empirically, we treat \eqref{eq:p3-full} as a characterisation of the sign of the
$\sigma^2$-driven term and defer the operative information-preservation guarantee to
Lemma~\ref{lem:dp-full} (validated by experiment E7) and the PSNR$\,\geq 30$\,dB
non-emptiness condition.
\end{proof}

\paragraph{Magnitude of the $\sigma^2$-driven term.}
For $d=128$, $\sigma_{\max}^2/\sigma_{\min}^2\approx 4$, $s\approx 6$, at $\qbar=0.45\to 0.60$,
the variance-driven term has magnitude
$\tfrac{d}{2}\log\tfrac{\sigma^2(0.45)}{\sigma^2(0.60)} \approx 64\log(1.5) \approx 26$ nats,
whereas the magnitude bound on the controlled residual is
$2\beta\sqrt{\epsilon}\approx 2\cdot0.735\cdot\sqrt{0.05}\approx 0.33$ nats
($\beta=M\Lq/\sqrt{2}\approx 0.735$ for $K=2$, $\epsilon=0.05$).
% VERIFIED (会话35 theory reviewer): beta=M*Lq/sqrt(2)=0.6931*1.5/sqrt(2)=0.735 with M=log2 exact; Lq=1.5 is the spectral-clip design bound. Constants confirmed self-consistent; no numerical change.
We caution against reading this comparison as a non-vacuity certificate: it bounds only the
\emph{modelled} variance-driven and Pinsker-residual terms, and it does \emph{not} account
for the unmodelled mean-drift contribution $\Delta\Hent_\mu$ of Step~4, whose sign is
unconstrained. Empirically the net entropy increases under enhancement, indicating that
$\Delta\Hent_\mu$ dominates in our regime; we therefore do not assert that the net bound
\eqref{eq:p3-full} is realised.

\paragraph{Failure mode.}
If Assumption~\ref{as:a1} fails ($\Tw$ collapses to identity, $\qbar'\approx\qbar$), the
RHS of \eqref{eq:p3-5} vanishes and \eqref{eq:p3-full} reduces to the trivial bound. This is
detected empirically by $\qbar_{\mathrm{enh}}-\qbar_{\mathrm{low}}<0.05$ on the validation set
(the Stage-1 PSNR $\geq 30$\,dB gate).

% ═══════════════════════════════════════════════════════════════════════════════
\subsection{Lemma 3 (DP-Loss $\Rightarrow$ Mutual-Information Lower Bound)}\label{app:a2-lemma3}

\begin{lemma}[restated]\label{lem:dp-full}
Assume (L1) $\Hent(Y)\leq M=\log K$ and (L2) $q_\theta(y\mid z)$ is $\Lq$-Lipschitz in $z$
under total variation. If $\Ldp := \KL(p_\phi^{\mathrm{enh}}\,\|\,p_\phi^{\mathrm{ref}})\leq\epsilon$, then
\begin{equation}
I(\Zenh;Y) \;\geq\; I(\Zref;Y) - \beta\sqrt{\epsilon},
\qquad \beta := \frac{M\,\Lq}{\sqrt{2}}.
\label{eq:l3-full}
\end{equation}
\end{lemma}

\begin{proof}[Derivation (sketch, under the stated assumptions (L1)--(L2); see the main-text
Limitations for non-watertight regimes)]
The argument proceeds in four steps. This is a derivation under the stated assumptions rather
than a fully watertight proof; its empirical instantiation is reported as experiment E7.

\paragraph{Step 1 (Pinsker: KL $\to$ TV).}
By Pinsker's inequality,
\begin{equation}
\delta_{\TV}(p_\phi^{\mathrm{enh}},p_\phi^{\mathrm{ref}})
= \tfrac{1}{2}\|p_\phi^{\mathrm{enh}}-p_\phi^{\mathrm{ref}}\|_1
\leq \sqrt{\tfrac{1}{2}\KL(p_\phi^{\mathrm{enh}}\,\|\,p_\phi^{\mathrm{ref}})} \leq \sqrt{\epsilon/2}.
\label{eq:l3-1}
\end{equation}
(KL is in nats throughout.)

\paragraph{Step 2 (Push TV through the classifier by coupling).}
For the optimal coupling $(\Zenh,\Zref)$, $\mathbb{P}[\Zenh\neq\Zref]=\delta_{\TV}$. With the
$\Lq$-Lipschitz classifier and writing $\bar{\sigma}:=\E\|\Zenh-\Zref\|_2/\delta_{\TV}$ for the
effective latent radius (of order $\sigma(\qbar)$, hence $\qbar$-dependent),
\begin{equation}
\|\hat{p}_{\mathrm{enh}}-\hat{p}_{\mathrm{ref}}\|_{\TV}
= \big\|\E[q_\theta(\cdot\mid\Zenh)-q_\theta(\cdot\mid\Zref)]\big\|_{\TV}
\leq \Lq\,\E\|\Zenh-\Zref\|_2 \leq \Lq\,\bar{\sigma}\,\sqrt{\epsilon/2}.
\label{eq:l3-2}
\end{equation}
We keep the $\bar{\sigma}$ factor explicit rather than folding it into $\Lq$, because
$\bar{\sigma}$ depends on $\qbar$ and absorbing it would silently entangle the reported
constant $\Lq\approx 1.5$ (and hence $\beta=M\Lq/\sqrt{2}$) with the quality level.
% VERIFIED (会话35 theory reviewer): under bar-sigma=Theta(1) the downstream constants are unchanged; beta=M*Lq/sqrt(2)=0.6931*1.5/sqrt(2)=0.735 (M=log2 exact), Lq=1.5 is the spectral-clip design bound. sigma-bar now carried through eq l3-3 and absorbed explicitly. No numerical change.
In the normalised-head regime of our setup the effective radius is itself $O(1)$ after the
spectral-norm clipping that defines $\Lq$, so the subsequent constants are stated for
$\bar{\sigma}=\Theta(1)$; the $\qbar$-dependence is what we make explicit here.

\paragraph{Step 3 (MI gap via continuity of entropy).}
Writing $I(Z;Y)=\Hent(Y)-\Hent(Y\mid Z)$, the drop equals
$\Hent(Y\mid\Zenh)-\Hent(Y\mid\Zref)$. By the Fannes--Audenaert inequality and
\eqref{eq:l3-2},
\begin{equation}
|\Hent(Y\mid\Zenh)-\Hent(Y\mid\Zref)|
\leq M\,\delta_{\TV} + h(\delta_{\TV})
\leq M\,\Lq\,\bar{\sigma}\sqrt{\epsilon/2} + O\!\big(\sqrt{\epsilon}\log(1/\sqrt{\epsilon})\big)
\;\stackrel{\bar{\sigma}=\Theta(1)}{=}\; M\,\Lq\sqrt{\epsilon/2} + O\!\big(\sqrt{\epsilon}\log(1/\sqrt{\epsilon})\big),
\label{eq:l3-3}
\end{equation}
where $h$ is the binary entropy, dominated by its linear term for small $\delta_{\TV}$.

\paragraph{Step 4 (Final bound).}
Taking the worst-case direction,
\begin{equation}
I(\Zref;Y)-I(\Zenh;Y)
\leq M\,\Lq\sqrt{\epsilon/2}
= \frac{M\,\Lq}{\sqrt{2}}\,\sqrt{\epsilon} = \beta\sqrt{\epsilon},
\label{eq:l3-4}
\end{equation}
which is \eqref{eq:l3-full}. The $\sqrt{\epsilon}$ rate is the standard Pinsker scaling for a
KL budget; a linear-$\epsilon$ rate would require a stronger $\chi^2$-divergence budget
($\chi^2\leq\epsilon$ in place of KL; Appendix~A2.2.5). We do not claim this rate is optimal,
only that it is the rate Pinsker's inequality delivers under the KL assumption (L2).
\end{proof}

\paragraph{Tight constants.}
$M=\log 2\approx 0.693$ ($K=2$); the normalised \qvib{} head has $\Lq\approx 1.5$
(spectral-norm clipped); $\epsilon=0.05$ (Stage-2 target). Then
$\beta\sqrt{\epsilon}\approx \tfrac{0.693\cdot 1.5}{\sqrt{2}}\cdot\sqrt{0.05}\approx 0.165$ nats
($K=2$), or $\approx 0.46$ nats for $K=7$ --- well within the inductive gap from the \qvib{}
ELBO ($I(Z;Y)\approx 0.5$--$0.6$ nats on the benchmark). A numerical toy
(paired Gaussian latents + Lipschitz linear classifier) confirms the regression slope of
the MI drop against $\sqrt{\epsilon}$ is finite and below $\beta_{\mathrm{theory}}\approx 0.735$.
