% =====================================================================
% §4 Diagnosis-Preserving Enhancement (C1 — hardest main leg). SHORT (9-page) version.
% Serves Claim C1; ACCEPTANCE C1.1-C1.5. caveman OFF.
% Anti-drift locks per STORY: deterministic restoration not generative; FiLM positioned for
%   diagnostic fidelity not pixel quality; completeness partially recoverable (R1);
%   Prop3 carried by E7 plus non-vacuity; derive/sketch wording (R2). Full proofs -> Appendix A2.
% Macros assumed in preamble: \visienhance \Tw \qbar \Ldp \KL \Zenh \Zref \tauenh \tauhigh
% =====================================================================

\section{Diagnosis-Preserving Enhancement}\label{sec:enhance}

Of the five quality dimensions, four admit deterministic restoration: sharpness, brightness, colour
temperature, and contrast. Completeness $q_3$ measures a truncated field of view and is only
\emph{partially recoverable}. Enhancement cannot synthesise tissue that was never captured, so the agent
treats low completeness as a query-for-retake signal and not as an enhancement target. This section
formalises the boundary of which dimensions are recoverable and within what severity window, and
\cref{sec:exp} measures it.

\subsection{Architecture: NAFNet with quality-conditional FiLM}\label{sec:arch}
\visienhance{} is a NAFNet backbone \citep{chen2022simple} with FiLM modulation \citep{perez2018film}
conditioned on the per-input quality vector, so restoration strength adapts to \emph{how} an image is
degraded instead of applying a fixed correction. The restoration map $\Tw(x,q)$ is
\emph{deterministic}. We avoid generative restoration backbones such as diffusion and adversarial
models, because such models can hallucinate diagnostic structure and invent a pigment network or
vascular pattern that was never present, which is clinically unacceptable. \cref{sec:discussion}
analyses why diffusion fails here. FiLM is \emph{not} a fidelity mechanism. Its empirical role under
ablation E8 is quality-conditional \emph{diagnostic} fidelity and not perceptual sharpening.

\subsection{DP-Loss: a feature-level diagnosis-preserving objective}\label{sec:dploss}
DP-Loss is the core contribution of this work. It is a feature-level training objective
that makes diagnostic fidelity the explicitly preserved quantity instead of pixel or perceptual
fidelity, and Lemma~\ref{lem:dp} in Section~\ref{sec:prop3lemma3} bounds its information loss.
Where prior restoration optimises perception with no guarantee that diagnostic content survives,
this objective constrains the enhancer to leave the diagnostic posterior intact. Its ablation E7
attributes the diagnosis-preserving effect to the design and not to restoration capacity. Training
minimises
\begin{equation}
\mathcal{L}=\mathcal{L}_1
  +\lambda_{\mathrm{LPIPS}}\,\mathcal{L}_{\mathrm{LPIPS}}
  +\lambda_{\mathrm{DP}}\,\Ldp
  +\lambda_{\mathrm{q}}\,\mathcal{L}_{\mathrm{hinge}},
\label{eq:total-loss}
\end{equation}
where $\mathcal{L}_1$ and $\mathcal{L}_{\mathrm{LPIPS}}$ govern pixel and perceptual fidelity, and the
diagnosis-preserving term is a \emph{feature-level} KL,
\begin{equation}
\Ldp \;=\; \KL\!\big(p_\phi^{\mathrm{enh}} \,\big\|\, p_\phi^{\mathrm{ref}}\big),
\label{eq:dploss}
\end{equation}
computed in the latent space of a frozen diagnostic feature extractor, the $1536\times7\times7$
penultimate features of an independent EfficientNet-B3 probe. $\Ldp$ aligns the enhanced image with
its clean reference in the space that drives the diagnosis and not in pixel space. A restorer can be
pixel-faithful yet still shift the diagnostic posterior, and \eqref{eq:dploss} penalises exactly that
movement. The quality hinge $\mathcal{L}_{\mathrm{hinge}}$ encourages $\qbar(\Tw(x,q))>\qbar(x)$ within
the salvage band, tying restoration to a measured quality gain.

\subsection{Mutual-information preservation}\label{sec:prop3lemma3}
% Lemma stated in body; directional entropy Proposition (prop:entropy) deferred to Appendix A2. R2: derive/sketch, not "prove".
The theoretical leg of DP-Loss connects the objective to diagnosis through a feature-level
mutual-information lower bound, stated here and derived in full in Appendix~\ref{app:a2}.
\input{appendix/A2_prop3_lemma3_compact}
Lemma~\ref{lem:dp} caps how much diagnostic information enhancement can destroy once the
feature-level KL is small. This bound carries the information-preservation role, and experiment E7
validates it empirically together with the Stage-1 PSNR$\,\geq30$\,dB non-vacuity condition. A
companion directional entropy result, Proposition~\ref{prop:entropy} in Appendix~\ref{app:a2},
characterises only the sign of one variance term and is \emph{not} a net entropy reduction. A
quality-conditioned probe finds that predictive entropy does not fall after enhancement. We therefore
state these results as derivations under the listed assumptions and not as watertight proofs; the full
treatment is in \cref{app:a2}.

\subsection{Three-stage training protocol}\label{sec:training}
We train $\visienhance$ in three stages, each isolating one term in \eqref{eq:total-loss}. Stage
\textbf{(1)} is pixel restoration with $\mathcal{L}_1$ and LPIPS up to the PSNR gate that guards the
non-vacuity condition of \eqref{eq:lemma3}. Stage \textbf{(2)} is DP-Loss fine-tuning that tightens
$\Ldp$ toward the budget $\epsilon$, and this stage is ablated as E7. Stage \textbf{(3)} is a
quality-hinge stage that calibrates the $\qbar$ gain inside the salvage band $[\tauenh,\tauhigh]$ and
credits enhancement only where it measurably raises quality.
