% =====================================================================
% §5 Query-for-Retake Agent (C2 — exclusive hook). ICLR S version (≤1.0 p).
% Compressed from drafts/s5_c2_agent.tex (1616w → ≤550w). caveman OFF.
% Anti-drift locks (STORY): retake = re-acquisition (collection loop), NOT abstain/defer (rule 8);
%   agent never diagnoses, only routes (C2); Thm2 LOCAL per-band only, deny global dominance,
%   honest negative → §7.7 (R11/R3); derive/under-assumption wording (R2); any per-input quality
%   scalar (R7); quality-conditioned bottleneck = footnote, off-the-shelf uncertainty (R13).
% Preserved labels (cited elsewhere): sec:agent, sec:agent-routing, sec:agent-thm2,
%   thm:agent-compact, eq:agent-policy-body. Cross-refs to app:a5 / app:a2-thm2 / sec:exp-dca /
%   sec:exp-boundary / sec:enhance kept. NUMBERS搬自源稿/ACCEPTANCE逐字, 未重算.
% =====================================================================

\section{Query-for-Retake Agent}\label{sec:agent}

The restorer of \cref{sec:enhance} is only safe to apply \emph{within} the salvage band. Outside it,
enhancement either does nothing useful or harms the diagnosis, as our boundary evidence in
\cref{sec:exp-boundary} shows. The agent therefore does not diagnose. From a per-input quality reading
it \emph{decides} whether to diagnose directly, to enhance and then diagnose, or to ask for a better
image. The last channel is the mechanistic contribution of this section. Existing reliability
mechanisms reviewed in \cref{sec:related}, namely selective prediction, learning-to-defer, and the
agentic pipeline AT-CXR, stop the moment an input is judged unsafe and abstain or defer without acting
on the cause of the failure, which is a recoverably bad acquisition. The query-for-retake channel
instead closes a \emph{collection} loop back to the sensor and requests a better image instead of
abstaining on or deferring the present one.

\paragraph{Four-channel routing.}\label{sec:agent-routing}
The policy $\pi^{*}$ reads a per-input quality scalar $\qbar(x)$, learned here by \visiscore{} but in
principle any per-input quality estimate such as a no-reference IQA score or a known corruption
severity, together with a predictive-entropy gate $\Hent(\hat p_x)\!\le\!\hhigh$, and selects one of
four actions $\{\actd,\acte,\actq,\actr\}$:
\begin{equation}
\pi^{*}(x)=
\begin{cases}
\actd & \qbar(x)\geq\tauhigh \quad\text{(direct diagnosis)},\\[1pt]
\acte & \qbar(x)\in[\tauenh,\tauhigh],\ \Hent(\hat p_x)\leq\hhigh \quad\text{(enhance-then-diagnose)},\\[1pt]
\actq & \taulow\leq\qbar(x)<\tauenh \quad\text{(query-for-retake)},\\[1pt]
\actr & \qbar(x)<\taulow\ \text{or}\ \big(\qbar(x)\in[\tauenh,\tauhigh],\ \Hent(\hat p_x)>\hhigh\big) \quad\text{(cautioned)},
\end{cases}
\label{eq:agent-policy-body}
\end{equation}
with grid-searched thresholds $\taulow\!=\!0.25<\tauenh\!=\!0.35<\tauhigh\!=\!0.50$. The entropy gate
separates a \emph{cautioned} high-quality case, one that is quality-sufficient yet intrinsically
ambiguous and beyond what enhancement can salvage, from a case the enhancement pathway can help. The
\textbf{query-for-retake} channel makes this mechanism concrete. Instead of abstaining
\citep{geifman2017selective} or deferring the case, the agent issues a \emph{re-acquisition request}
that closes a collection loop around the diagnostic model. Image-quality work that flags re-capture
\citep{vodrahalli2021teledermatology,jalaboi2023explainable} decouples the flag from the diagnostic
objective and offers no risk analysis. Here the re-acquisition decision is made jointly with the
diagnostic risk and embedded in the risk-bounded routing policy below. To our knowledge, this is the
first \emph{risk-bounded re-acquisition loop coupled jointly with the diagnostic decision} in this
setting, a scope we delimit deliberately. The agent never emits a diagnosis on its own. Each channel
hands the original, restored, or re-acquired image to the same downstream classifier. The uncertainty
signal is taken off the shelf, so routing does not depend on any bespoke quality-conditioned
bottleneck.\footnote{A quality-conditioned bottleneck variant was evaluated as an uncertainty source
but showed no measurable advantage over the standard variant here.} Implementation is a lightweight
ReAct-style controller backed by a small language model with a deterministic rule fallback applying
\eqref{eq:agent-policy-body}, so routing is never less safe than the rule. The prompt, the fallback
specification, and a worked trace are in Appendix~\ref{app:a5}. The retake channel fires where it
should and rises monotonically as quality falls across the high-quality, moderate, and severe bands,
as reported in \cref{sec:exp-boundary}.

\paragraph{A local, per-band risk guarantee (Theorem~\ref{thm:agent-compact}).}\label{sec:agent-thm2}
\emph{Within} the salvage band, and under a diagnosis-preserving restorer satisfying the information
lower bound of \cref{sec:enhance}, we derive that enhance-then-diagnose attains a strictly lower
conditional expected risk than direct diagnosis,
$\Renh(x)\le\Rdirect(x)-\Delta(\qbar(x),\Tw)$ with $\Delta(\qbar,\Tw)>0$ for
$\qbar\in[\tauenh,\tauhigh]$. The per-band reductions are realised empirically with bootstrap $95\%$
CIs excluding zero, as reported in \cref{sec:exp-boundary}. The decision-theoretic setup, the auxiliary
lemmas, and the finite-sample companion are in Appendix~\ref{app:a2-thm2}.

\input{appendix/A2_3_theorem2_compact}

This is deliberately a \emph{local} result. It certifies enhancement as the risk-reducing action
exactly where the band holds and is silent elsewhere. \textbf{We make no claim that the routed agent
dominates direct diagnosis globally.} The within-band gain does not aggregate to a global ordering,
because the query channel's safety relies on a re-acquisition operator that improves quality only in
expectation, and the thresholds are fixed by grid search. Consistently, our triage simulation reports
an honest \emph{negative} global result in \cref{sec:exp-dca}. At a fixed referral budget, direct
triage attains the higher auto-handled sensitivity, and net benefit is statistically indistinguishable
across variants. We therefore treat the global utility of routing as an open and empirically negative
question, and we leave \emph{learning} the thresholds to future work. The band-local statement is
moreover a population result over the salvage-band subpopulation. It does \emph{not} certify a
per-class reduction, and the malignant subgroup stays in the query-for-retake regime instead of the
enhance regime; its net-negative salvage is our hardest boundary evidence, reported in
\cref{sec:exp-boundary}.
% TODO(verifier): small-LM identity kept generic pre-deanonymisation (see internal STORY).
% All numbers (0.818/0.788, retake 0.889/0.651/0.055, TAU 0.25/0.35/0.50)搬自源稿/ACCEPTANCE, 未重算.
