% ============================================================
% §6  Experimental Setup  (S version, 9-page short).
% Compressed from drafts/s6_setup.tex (716 w -> <=300 w).
% Drift contract: serves C1 (diagnosis-preserving enhancement) lever; this
% section is the evaluation substrate, NOT a contribution (benchmark demoted).
% Full benchmark construction -> Appendix A4.
% R-rules kept: 6 enhancement baselines, diffusion forbidden (R8), bootstrap
% 2000-sample 95% CI (R6), BMVC cited not reused (R10), anonymise (R4).
% Numbers copied verbatim from drafts/s6_setup.tex; none recomputed.
% ============================================================

\section{Experimental Setup}
\label{sec:setup}

\paragraph{Datasets and metrics.}
We evaluate on dermoscopic skin-lesion imagery from the public ISIC-2020
collection under a fixed train/validation/test split. Paired diagnostic
evaluations use the held-out paired subset of $n=3627$ images with $117$
melanoma-positive cases, together with a frozen EfficientNet-B3 classifier that is
held fixed so that any change in diagnostic AUC is attributable to the input and not
to retraining. We report PSNR, SSIM, the frozen-B3 AUC and induced $\Delta\mathrm{AUC}$,
the enhanced-vs-reference agreement rate, and the DP-Loss feature-level KL. Every
such figure carries a bootstrap $2000$-sample $95\%$ confidence interval with a
McNemar test on paired comparisons. Quality is made a known variable through
controlled degradations along five physically parameterised dimensions at a
\emph{moderate} tier: blur, brightness, completeness, colour shift, and contrast. The
split details, the severity grid, and the full metric and degradation protocol are
deferred to Appendix~\ref{app:a4}. A companion calibration study is cited
\citep{ANON-BMVC} but not reused.

\paragraph{Enhancement baselines.}
We compare against six publicly available deterministic restoration backbones
spanning convolutional and transformer designs: Restormer~\citep{zamir2022restormer},
NAFNet~\citep{chen2022simple}, MIRNet-v2~\citep{zamir2022mirnetv2},
SwinIR~\citep{liang2021swinir}, Uformer-B~\citep{wang2022uformer}, and
Real-ESRGAN~\citep{wang2021realesrgan}, each run zero-shot from official weights
under the identical protocol and frozen B3, and additionally fine-tuned on the same
degraded split under a matched protocol detailed in Appendix~\ref{app:a4-ft}.\todo{R1 ft 协议}
We deliberately exclude diffusion and other generative restorers. On dermoscopy they
can hallucinate diagnostically salient structure such as invented pigment networks or
vessels. This is a medical-safety hazard that we analyse in \S\ref{sec:discussion} and
do not admit as a baseline.

