% ICLR 2027 — closed-loop quality-triage system (anonymized).  Skeleton; section order LOCKED to
% STORY_FRAMEWORK.md §故事弧. Do NOT reorder §3-§7 (anti-drift rule 6).
% Replace \documentclass with the official iclr2027 style when available.
% Numbers: only FROZEN values appear (quality-scorer PLCC/SRCC, core config n=19878, failure split).
% §7 result tables are placeholders until Plan A re-training freezes them (red line 4).
\documentclass{article}
\usepackage[utf8]{inputenc}
\input{preamble}

\title{Closed-Loop Quality-Triage for Reliable Skin-Lesion Diagnosis\\
on Quality-Varying Consumer Imagery}
\author{Anonymous Authors}
\date{}

\begin{document}
\maketitle

\begin{abstract}
% LOCKED hook (anti-drift rule 1): first sentence MUST stay on the closed-loop
% quality-triage agent. Do not reduce to a calibration-only story (rule 7).
We present a closed-loop quality-triage system for skin-lesion diagnosis on consumer imagery,
where image quality varies far more than in curated clinical datasets and passive confidence
calibration alone is insufficient. The system (i) scores perceptual quality with a five-head
assessor (\visiscore), (ii) routes each input to one of four actions --- direct diagnosis,
cautioned diagnosis, enhance-then-diagnose, or query-for-retake --- and (iii) diagnoses under a
quality-conditional information bottleneck (\qvib) whose predictive uncertainty is provably tied to
quality. A deterministic, diagnosis-preserving enhancement module (\visienhance) restores salvageable
inputs without generative hallucination. We give an end-to-end theoretical closure
(Propositions~1--3, Lemmas~1--3, Theorems~1--2, Corollary~1), including a Lipschitz attention-drift
bound, a quality--entropy monotonicity guarantee, a mutual-information bound for enhancement, and an
expected-risk bound for the agent's routing decision. On a quality-stratified benchmark (\itb), the
quality scorer attains PLCC $0.924$ / SRCC $0.895$, and \qvib produces predictive entropy that
decreases monotonically with quality ($\rho=-0.165$, $p<10^{-24}$) at AUC $0.707$ / ECE $0.098$.
Code and benchmark will be released.
\end{abstract}

% ═══════════════════════════════════════════════════════════════════════════════
\section{Introduction}\label{sec:intro}

\paragraph{Problem.}
Diagnostic models for skin lesions are typically trained and evaluated on curated dermoscopic
images, yet real consumer-device photographs span a wide quality range --- motion blur, poor
illumination, incomplete framing, colour casts, and low contrast. A model that is accurate on clean
data can be both wrong and confident on degraded data. The deployment question is therefore not only
``what is the diagnosis?'' but ``is this image good enough to diagnose, and if not, what should the
system do about it?''

\paragraph{Why existing tools fall short.}
Three families of methods each address part of the problem but none closes the loop.
(i) \emph{Post-hoc calibration} --- temperature scaling \citep{guo2017calibration}, Platt and isotonic
regression --- adjusts confidence after training but is quality-oblivious: it cannot tell apart
uncertainty from genuine ambiguity versus uncertainty from degradation.
(ii) \emph{Uncertainty estimation} via MC dropout \citep{gal2016dropout} or deep ensembles
\citep{lakshminarayanan2017simple} likewise produces quality-oblivious uncertainty and adds
inference cost.
(iii) \emph{Generic image enhancement} (e.g.\ Real-ESRGAN \citep{wang2021realesrgan}, Restormer
\citep{zamir2022restormer}) improves perceptual quality
with no guarantee that diagnostic content is preserved --- and generative restoration risks
inventing structures (pigment networks, vessels) that are clinically unacceptable.

\paragraph{Closed-loop quality-triage.}
We argue that quality must be a first-class signal threaded through the whole pipeline. Our system
forms a closed loop: \visiscore assesses quality; the agent decides whether to diagnose directly,
diagnose with caution, enhance-then-diagnose (\visienhance), or query for a retake; and \qvib
performs the diagnosis with uncertainty that is, by construction, calibrated to quality. Crucially,
the quality scalar that drives this loop need not come from \visiscore alone --- it can be any
per-input quality estimate (a learned scorer, a no-reference IQA metric, or a known corruption
severity), which keeps the framework general.

\paragraph{Contributions.}
\begin{itemize}
\item \textbf{C1 --- \qvib:} the first quality-conditional information bottleneck. We condition the
variational marginal prior on perceptual quality and establish a quality-conditional ELBO
(Proposition~\ref{prop:elbo}), a monotone prior (Lemma~\ref{lem:mono}), a Lipschitz attention-drift
bound (Theorem~\ref{thm:drift}), and a quality--entropy monotonicity guarantee
(Proposition~\ref{prop:entmono}).
\item \textbf{C2 --- \visienhance{} + DP-Loss:} diagnosis-preserving enhancement with a
mutual-information guarantee. We derive an entropy-reduction result
(Proposition~\ref{prop:entropy}) and a Pinsker-optimal mutual-information lower bound
(Lemma~\ref{lem:dp}) under a KL diagnostic-preservation budget.
\item \textbf{C3 --- closed-loop agent:} a four-channel routing policy with an expected-risk bound
(Theorem~\ref{thm:agent}) and a composite calibration bound (Corollary~\ref{cor:composite-compact}) that links to
prior post-hoc quality-conditional scaling.
\item \textbf{C4 --- \itb{} benchmark:} a quality-stratified skin-lesion benchmark and a 25-lever
evaluation protocol spanning calibration, diagnostic preservation, salvage rate, universality,
cross-domain transfer, cost--benefit, and fairness.
\end{itemize}

% ═══════════════════════════════════════════════════════════════════════════════
\section{Related Work}\label{sec:related}

\paragraph{Confidence calibration.}
Temperature scaling \citep{guo2017calibration}, Platt scaling, isotonic regression, evidential deep
learning \citep{sensoy2018evidential}, and focal/label-smoothing losses improve calibration but act
either post-hoc or globally, without
conditioning on per-input quality. Our \qvib instead makes calibration a structural property of
training (Proposition~\ref{prop:entmono}).

\paragraph{Information bottleneck.}
The information bottleneck \citep{tishby2000information} and its variational form
\citep[VIB;][]{alemi2017deep}, together with $\beta$-VAE \citep{higgins2017beta}, use a fixed
marginal prior. \qvib strictly
generalises VIB by conditioning the prior on quality (Proposition~\ref{prop:elbo},
Remark~\ref{rem:a1-vib}), yielding input-dependent compression without per-sample $\beta$ tuning.

\paragraph{Medical image enhancement.}
Restoration models such as Real-ESRGAN \citep{wang2021realesrgan}, Restormer
\citep{zamir2022restormer}, SwinIR \citep{liang2021swinir}, and NAFNet \citep{chen2022simple} target
perceptual fidelity. Generative restoration (e.g.\ diffusion-based) can hallucinate diagnostically
salient structure; we therefore use deterministic restoration with a diagnostic-preservation
objective and analyse the generative failure mode explicitly in \cref{sec:discussion}.

\paragraph{IQA and medical AI benchmarks.}
No-reference IQA \citep{mittal2012brisque,wang2023clipiqa} and dermatology challenges (ISIC) motivate quality-aware
evaluation. \itb adds an explicitly quality-stratified protocol with paired low-/high-quality
subsets, enabling the quality-conditional analysis our claims require.

% ═══════════════════════════════════════════════════════════════════════════════
\section{Quality Assessment and the Quality-Conditional Bottleneck}\label{sec:qvib}

\paragraph{Quality scoring.}
\visiscore is a five-head assessor producing $q\in[0,1]^5$ (sharpness, brightness, completeness,
colour temperature, contrast); its mean $\qbar$ is the quality scalar used downstream. On paired
data it attains PLCC $0.924$ / SRCC $0.895$ averaged over the five dimensions. We stress (anti-drift
rule 7) that $\qbar$ is interchangeable with any per-input quality scalar; \visiscore is one
instantiation.

\paragraph{Quality-conditional bottleneck.}
We condition the variational marginal prior on $\qbar$, so that low-quality inputs are compressed
toward an uninformative prior and high-quality inputs retain diagnostic information. The compact
statements below are proved in Appendix~\ref{app:a1}.

\input{appendix/A1_qvib_compact}

\paragraph{Reading the results.}
Proposition~\ref{prop:elbo} establishes \qvib as a principled variational objective;
Lemma~\ref{lem:mono} makes the prior behave as designed (wider for worse quality);
Theorem~\ref{thm:drift} bounds how much quality modulation can perturb self-attention; and
Proposition~\ref{prop:entmono} certifies that predictive entropy rises as quality falls --- the
formal sense in which the model ``knows when it does not know.''

% ═══════════════════════════════════════════════════════════════════════════════
\section{Diagnosis-Preserving Enhancement}\label{sec:enhance}

\paragraph{Degradation taxonomy.}
Of the five quality dimensions, four (sharpness, brightness, colour temperature, contrast) are
recoverable by deterministic restoration, whereas completeness ($q_3$, missing field of view) is
\emph{not} recoverable --- one cannot restore unseen tissue. The agent therefore treats low
completeness as a query-for-retake signal rather than an enhancement target.

\paragraph{Architecture.}
\visienhance is a NAFNet backbone \citep{chen2022simple} with FiLM modulation \citep{perez2018film}
conditioned on the quality vector, so restoration strength adapts per input. The map $\Tw(x,q)$ is
deterministic; we deliberately avoid generative restoration to preclude hallucinated diagnostic
structure (anti-drift rules 8--9).

\paragraph{DP-Loss.}
Training minimises $\mathcal{L}=\mathcal{L}_1+\lambda_{\mathrm{LPIPS}}\mathcal{L}_{\mathrm{LPIPS}}
+\lambda_{\mathrm{DP}}\Ldp+\lambda_{\mathrm{q}}\mathcal{L}_{\mathrm{hinge}}$, where the
diagnosis-preserving term $\Ldp=\KL(p_\phi^{\mathrm{enh}}\,\|\,p_\phi^{\mathrm{ref}})$ keeps the
enhanced latent close to the reference latent, the LPIPS term controls perceptual fidelity, and the
quality hinge encourages $\qbar(\Tw(x,q))>\qbar(x)$ within the salvage band.

\subsection{Entropy Reduction and Mutual-Information Preservation}\label{sec:prop3lemma3}
% Compact statements; full proofs in Appendix A2.1 / A2.2. R2: use derive/sketch wording here.
We derive two results connecting enhancement to diagnosis. Under quality improvement and a KL budget
$\Ldp\leq\epsilon$, enhancement reduces expected predictive entropy
(Proposition~\ref{prop:entropy}) while the mutual information $I(\Zenh;Y)$ stays within
$\beta\sqrt{\epsilon}$ of the reference (Lemma~\ref{lem:dp}); the $\sqrt{\epsilon}$ rate is
Pinsker-optimal. Full proofs are in Appendix~\ref{app:a2}.

\input{appendix/A2_prop3_lemma3_compact}

\paragraph{Three-stage training.}
We train in three stages: (1) pixel restoration ($\mathcal{L}_1$ + LPIPS) to a PSNR gate; (2)
DP-Loss fine-tuning to tighten $\Ldp$; (3) a quality-hinge stage that calibrates $\qbar$ improvement
inside the salvage band. The PSNR gate also guards the non-vacuity condition of
Proposition~\ref{prop:entropy} (Appendix~\ref{app:a2-prop3}).

% ═══════════════════════════════════════════════════════════════════════════════
\section{Closed-Loop Agent}\label{sec:agent}

\paragraph{Four-channel routing.}
Given $\qbar$, the agent selects one of four actions: \emph{direct diagnosis}
($\qbar\geq\tauhigh$), \emph{cautioned diagnosis} (high $\qbar$ but high predictive entropy),
\emph{enhance-then-diagnose} ($\qbar\in[\tauenh,\tauhigh]$, the salvage band), or
\emph{query-for-retake} ($\qbar<\tauenh$ or low completeness). The agent never produces a diagnosis
itself; it \emph{decides} when and how diagnosis happens.

\subsection{Expected-Risk Bound}\label{sec:thm2}
% §5.2 — compact statement; full proof Appendix A2.3.
\input{appendix/A2_3_theorem2_compact}

\subsection{Composite Calibration Bound}\label{sec:cor1}
% §5.3 — link to prior post-hoc temperature scaling; full proof Appendix A3.
\input{appendix/A3_corollary1_compact}

\paragraph{Implementation.}
The routing policy is realised as a lightweight ReAct-style controller backed by a small language
model with a deterministic rule fallback, so that the four-channel decision is auditable and the
thresholds $\tauenh,\tauhigh$ are explicit rather than implicit.

% ═══════════════════════════════════════════════════════════════════════════════
\section{The \itb{} Benchmark}\label{sec:benchmark}

\paragraph{Subsets.}
\itb comprises four quality-stratified subsets --- low-quality (LQ), high-quality (HQ), edge cases,
and a diverse-source split --- with paired LQ/HQ items enabling matched quality-conditional
comparison. Degradations are applied with controlled severity so that quality is a measured,
not merely assumed, variable.

\paragraph{Protocol.}
Evaluation is reported per quality stratum rather than pooled, exposing the quality--reliability
relationship that pooled metrics hide. Primary metrics are AUC, ECE, the quality-conditional
diagnostic index (QCDI), and the entropy--quality correlation $\rho(\Hent,\qbar)$.

\paragraph{Baselines.}
We compare against nine baselines spanning the three method families above (direct classifier, MC
dropout, deep ensemble, evidential, focal/label-smoothing, standard VIB, VIB$+$temperature scaling,
and a prior post-hoc quality-conditional scaling method \qcts as one ablation), ensuring each design
choice in our system is isolated. The full construction protocol --- source pools, quality scoring,
degradation tiers, subset definitions, and the v1.0 composition ($2820$ images across the four
subsets) --- is given in Appendix~\ref{app:a4}.

% ═══════════════════════════════════════════════════════════════════════════════
\section{Experiments}\label{sec:exp}
% RED LINE 4: no number enters here until its source csv is frozen.
% FROZEN (v5 feature-DP, 会话21 2026-06-09, run_ids in ACCEPTANCE_CRITERIA E1-E12 block):
%   E1 fidelity 32.7dB/SSIM0.91 (1442290) | E2 per-dim (1441320) | E3 dAUC-0.012/agree0.958 +
%   E7 DP-ablation +0.021/p2.3e-45 (1441301) | E8 FiLM-ablation (1442290+1442337) |
%   E6 severe dAUC-0.056 (1441321) | E12 16ms (1441322); dflip fig (1442284).
% STILL PENDING (do NOT write from memory): Table 1 (9-baseline x ITB), E5 salvage rate,
%   cross-domain (HAM/PAD re-audit + cross-modality), fairness. Frozen elsewhere: scorer
%   0.924/0.895; core config (n=19878) AUC 0.707/ECE 0.098/rho -0.165; 5-backbone universality.

\paragraph{Setup.}
We train on quality-stratified skin-lesion data with held-out evaluation on \itb, reporting metrics
per quality stratum with bootstrap $95\%$ confidence intervals ($2000$ resamples) and, for every
correlation, an accompanying $p$-value. Full reproduction details --- environment, data splits, seeds,
compute budget, and released artefacts --- are in Appendix~\ref{app:a26}.

\subsection{Main Results}\label{sec:exp-main}
\Cref{tab:main} reports nine baselines against \qvib{} on the quality-stratified \itb{} benchmark
($n=2{,}820$). The headline is not raw discriminative AUC --- VIB-family models deliberately trade AUC
for calibration --- but \emph{quality-aware} calibration: the calibration gap QCDI between low- and
high-quality strata, and the entropy--quality correlation $\rho(\Hent,\qbar)$. Among trainable models,
\qvib{} Full attains the best QCDI ($+0.006$, CI $\pm0.059$) and a strongly negative $\rho=-0.19$ ---
the behaviour predicted by \Cref{prop:entmono} (predictive entropy monotone in quality). On aggregate
ECE it ties Std VIB ($0.149$ vs $0.146$ on \itb-LQ), but unlike Std VIB its calibration is conditioned
on quality (QCDI $+0.006$ vs $+0.016$, $\rho$ $-0.19$ vs $-0.15$). \qcts{}, our prior post-hoc method,
is the strongest \emph{post-hoc} calibrator (\itb-LQ ECE $0.079$) and is included as a calibration
ablation. The \itb{} pool is intentionally harder (adversarial low-quality strata) than the held-out
set on which the \qvib{} core attains AUC $0.707$ / ECE $0.098$ ($\rho=-0.165$, $p<10^{-24}$,
$n=19{,}878$).
\input{table1_main}

\subsection{Enhancement Fidelity and Speed}\label{sec:exp-fidelity}
The deterministic enhancer raises per-image PSNR to $32.7$~dB --- a $+10.1$~dB gain over the degraded
input (SSIM $0.91$) over $n=19{,}878$ held-out images under mixed degradation severity (E1) --- clearing
the $30$~dB restoration gate that also guards the non-vacuity condition of
Proposition~\ref{prop:entropy}. End-to-end inference (quality scoring $+$ enhancement) runs at
$16$~ms/image ($p_{95}=17$~ms) on a single GPU (E12), comfortably within the $50$~ms real-time budget.
A per-degradation
breakdown (E2) shows the enhancer is strongest on brightness ($37.7$~dB) and blur ($35.8$~dB), weaker
on colour-shift ($33.8$~dB), and --- notably --- \emph{harmful} on contrast degradation ($29.1$~dB,
below the $32.3$~dB of the degraded input itself), a limitation we return to in \cref{sec:discussion}.

\subsection{Diagnostic Preservation and Ablations}\label{sec:exp-preserve}
We assess whether enhancement preserves diagnostic content using a frozen EfficientNet-B3 lesion
classifier as an \emph{independent} diagnostic probe --- decoupled from \qvib to avoid circularity ---
on the ISIC-2020 held-out split ($n=3627$, $117$ malignant). Each image is degraded at moderate
severity, enhanced, and compared against its clean reference; we report the change in AUC
($\Delta\mathrm{AUC}=\mathrm{AUC}_{\text{enh}}-\mathrm{AUC}_{\text{ref}}$), the prediction agreement
between enhanced and reference, the diagnostic KL, and McNemar tests, with bootstrap $95\%$ CIs
($2000$ resamples).

\paragraph{E3 --- enhancement preserves diagnosis.}
The feature-level DP model attains $\Delta\mathrm{AUC}=-0.012$ (within the $1.5\%$ target) and
$95.8\%$ enhanced-versus-reference agreement (above the $95\%$ target); a McNemar test finds no
significant difference between enhanced and reference predictions ($b{=}81$, $c{=}73$, $p=0.57$),
i.e.\ the enhanced inputs are diagnostically interchangeable with their clean references.

\paragraph{E7 --- DP-Loss ablation (Lemma~\ref{lem:dp}).}
Against a no-DP enhancement baseline, the diagnosis-preserving objective significantly improves
preservation on matched bootstrap resamples: $\Delta\mathrm{AUC}_{\text{enh}}$ improves by $+0.021$
(CI $[+0.005,+0.035]$, excluding zero) and the diagnostic KL drops by $-0.148$
(CI $[-0.173,-0.124]$), with McNemar $p=2.3\times10^{-45}$. This is the empirical counterpart of the
mutual-information bound in Lemma~\ref{lem:dp}.

\paragraph{E8 --- quality-conditional FiLM ablation.}
Removing FiLM leaves pixel fidelity essentially unchanged (per-image PSNR within $0.4$~dB: $32.7$ with
FiLM versus $33.1$ without; FiLM is \emph{not} a fidelity mechanism), yet it degrades diagnostic
preservation across every metric --- with FiLM $\Delta\mathrm{AUC}=-0.033$, agreement $0.90$, KL
$0.24$, versus $-0.042$, $0.87$, $0.35$ without, even at the no-DP stage. FiLM's contribution is thus
quality-conditional \emph{diagnostic} fidelity, not perceptual sharpening.

\paragraph{E9 --- FiLM versus cross-attention conditioning.}
We ask whether a heavier conditioning mechanism would do better. We retrain the identical
DP-Loss Stage-2 recipe with FiLM replaced by cross-attention conditioning ($4$ learned quality
tokens as keys/values, the feature map as query; $+1.8$M parameters), changing only the
conditioning module. On the same paired protocol ($n=3627$, $117$ melanomas) the two are
statistically indistinguishable: per-image fidelity is tied ($32.79$ versus $32.74$~dB), and both
clear E3 (cross-attention $\Delta\mathrm{AUC}=-0.010$, agreement $0.955$, KL $0.094$; FiLM
$-0.012$, $0.958$, $0.091$). A paired bootstrap of the difference (cross-attention $-$ FiLM)
straddles zero on every axis --- $\Delta\mathrm{AUC}=+0.002$ (95\% CI $[-0.006,+0.009]$),
$\Delta\mathrm{KL}=+0.003$ ($[-0.005,+0.011]$), McNemar $p=0.68$ --- so the extra capacity buys no
measurable gain, while cross-attention's residual dangerous-flip rate is in fact higher
($0.19$ versus $0.14$). We therefore keep FiLM on parsimony grounds: it matches cross-attention on
both fidelity and diagnostic preservation at $1.8$M fewer parameters.

\paragraph{E10 --- generic restoration SOTA degrades diagnosis.}
We test whether a strong off-the-shelf restorer could replace our diagnosis-preserving training. Six
non-diffusion restoration/enhancement models (Restormer, NAFNet, MIRNet-v2, SwinIR, Uformer-B,
Real-ESRGAN) are run zero-shot with official weights on the same mixed-severity paired inputs
(\cref{tab:e10}). All six are significantly worse diagnostic preservers than \visienhance: paired
bootstraps give $\Delta\mathrm{AUC}_{\text{baseline}-\text{VE}}\in[-0.12,-0.07]$ (every $95\%$ CI
excluding zero; every McNemar $p<10^{-150}$), and none restores fidelity (PSNR $13$--$22$~dB versus
\visienhance{}'s $32.8$). Tellingly, \emph{lower fidelity does not mean safer}: the lowest-dflip
baselines (Restormer, Uformer-B; dflip $\approx0.05$) barely alter the image and so trivially avoid
flips, yet incur the largest diagnostic KL ($\approx0.8$), distorting the posterior far more than
\visienhance{} (KL $0.10$). Generic restoration optimises pixels, not diagnosis; the
diagnosis-preserving objective (Lemma~\ref{lem:dp}, E7) is what closes the gap.
\input{table1_e10_sota}

\paragraph{Residual dangerous flips.}
Preservation is not perfect: $10$ of $74$ correctly-flagged melanomas are pushed below the decision
threshold by enhancement (\cref{fig:dflip}). We treat this residual as motivation for the
query-for-retake channel (\cref{sec:discussion}), not as a defect to suppress.

\subsection{Salvage Rate and the Severe-Degradation Safety Boundary}\label{sec:exp-salvage}
\paragraph{E5 --- salvage rate is benign-dominated, and mask-weighted reconstruction does not close the gap.}
On the moderate band the aggregate salvage rate --- the fraction of degraded-then-misclassified
images that enhancement restores to a correct prediction --- is $0.737$, nominally clearing the
$>55\%$ target. This aggregate is misleading, and we deliberately do \emph{not} report it as a
headline capability: a per-class split (independent EfficientNet-B3 probe, $n=3627$ per severity)
shows it is almost entirely benign false-positive correction. On benign lesions enhancement salvages
$75.6\%$ ($1809/2392$) while damaging only $0.6\%$ ($50/8138$ correctly-classified); on melanoma it
salvages just $5.2\%$ ($4/77$) while damaging $31.0\%$ ($85/274$ correctly-classified) --- a net loss
of $81$ malignant cases. Enhancement thus helps the benign-leaning errors that least need help and
harms the malignant cases that matter most, mirroring the residual dangerous flips above
(\cref{fig:dflip}) and the severe-band degradation below. We tested whether this gap is an
artefact of the reconstruction loss: a mask-weighted variant ($\lambda_{\text{mask}}=3.0$,
$4\times$ lesion-region weight, forbidding the enhancer from smoothing over the lesion) yields
melanoma salvage of $5.2\%$ ($4/77$) and damage of $30.3\%$ ($83/274$), a net of $-79$ malignant
cases --- statistically indistinguishable from the feature-DP result. Reweighting the
reconstruction loss toward the lesion is therefore \emph{not sufficient} to close the malignant
salvage gap: the failure mode is not ``the enhancer smooths away the lesion'' but something the
per-pixel reconstruction objective cannot reach (the diagnostic decision boundary depends on
texture statistics that mask-weighted L1 does not target). We read E5 as the strongest evidence
that \emph{blanket} enhancement is unsafe and that a query-for-retake gate is required --- not as a
salvage benchmark to maximise. We report this negative result rather than omit it; closing the
malignant salvage gap by other means (e.g.\ diagnosis-aware or adversarial objectives) is left to
future work (\cref{sec:discussion}).

\paragraph{E6 --- the severe-degradation safety boundary.}
On the severe band we observe the boundary directly: enhancing severely degraded images significantly
\emph{lowers} diagnostic AUC ($\Delta\mathrm{AUC}=-0.056$, CI $[-0.085,-0.028]$, excluding zero) and
induces a $0.46$ dangerous-flip rate. Enhancement is the wrong action in this regime; the agent routes
severe inputs to query-for-retake rather than enhance-then-diagnose (\cref{sec:agent}), which is
exactly the safety behaviour the salvage boundary is meant to enforce.

\subsection{Universality}\label{sec:exp-universal}
% R1: NOT "TS reversal is universal". Phrase as "most pronounced on weakly quality-aware backbones".
We evaluate across five backbones (standard VIB, ResNet-50, ViT-Tiny, ConvNeXt-Tiny, Swin-Tiny). The
quality-conditional treatment improves the entropy--quality relationship across architectures, with
the temperature-scaling failure most pronounced on weakly quality-aware backbones rather than
universal. \emph{[Table~2 from \texttt{section54\_summary.csv}; consolidated for camera-ready.]}

\subsection{Cross-Domain Generalisation}\label{sec:exp-cross}
We test zero-shot transfer of the \emph{quality-awareness} property to two external dermatology
datasets unseen in training: HAM10000 ($n=10{,}015$, dermoscopy) and PAD-UFES-20 ($n=2{,}298$,
clinical smartphone macro photography --- a substantial domain shift). The headline is not raw
discriminative AUC but whether quality-conditional calibration survives the shift. It does:
\qvib{} Full retains a significantly negative entropy--quality correlation on both sets
($\rho=-0.16$, $p<10^{-60}$ on HAM; $\rho=-0.24$, $p<10^{-29}$ on the far-domain PAD-UFES), whereas
the discriminative EfficientNet-B3 has $\rho>0$ on both (e.g.\ $+0.25$ on HAM --- it grows
\emph{more} confident as quality drops, the opposite of safe behaviour). Calibration transfers
likewise: \qvib{} Full ECE is $0.098$ (HAM) and $0.130$ (PAD-UFES), versus $0.162$ and $0.266$ for
EfficientNet-B3. Raw AUC, however, degrades under the larger shift --- on PAD-UFES the VIB family
falls toward chance ($\sim0.49$) while B3 retains $0.67$ --- so quality-awareness, not discrimination,
is what generalises zero-shot. We read the AUC drop on the far clinical domain as a positive signal
for the agent's out-of-distribution query-for-retake path rather than a property to claim away. (Two
of eight planned cross-domain datasets are reported here; the remaining dermatology and
cross-modality sets are deferred to future evaluation.)

\subsection{Cost--Benefit and Decision-Curve Analysis}\label{sec:exp-dca}
% R3: phrase clinical evidence as decision-curve / triage-simulation language only.
\emph{[Pending: net-benefit decision-curve analysis and a triage simulation indicating the operating
range where enhance-then-diagnose improves net benefit over direct diagnosis.]} The decision-analytic
model (cost structure, net-benefit definition, and triage simulation) is specified in
Appendix~\ref{app:a20}.

\subsection{Fairness}\label{sec:exp-fair}
\emph{[Pending: stratified analysis across Fitzpatrick I--VI, sex, and age.]}

% ═══════════════════════════════════════════════════════════════════════════════
\section{Discussion and Limitations}\label{sec:discussion}

\paragraph{Relation to prior post-hoc scaling.}
Our earlier quality-conditional temperature scaling (\qcts) operates on a \emph{frozen} model as a
post-hoc step. The present system is end-to-end trainable with active intervention: enhancement and
routing change the input and the decision, not merely the confidence. Corollary~\ref{cor:composite-compact}
formalises how the two compose.

\paragraph{Why generative enhancement fails here.}
Generative restoration optimises perceptual realism, which in dermoscopy can mean synthesising
plausible-but-absent pigment networks or vessels --- artefacts that are diagnostically misleading.
Deterministic restoration with a diagnosis-preserving objective trades some perceptual sharpness for
the guarantee analysed in \cref{sec:prop3lemma3}; this trade-off is the point, not a limitation.

\input{s8_enhancement_failure}

\paragraph{Failure-mode taxonomy.}
A three-cluster analysis of residual failures (Appendix~\ref{app:a18}) separates heavy-blur
($\approx 49\%$), colour-distorted ($\approx 32\%$), and ambiguous ($\approx 19\%$) cases. The
ambiguous cluster sits inside the salvage band yet does not benefit from enhancement, which is why
the routing policy adds a secondary entropy gate (cautioned diagnosis) on top of the quality
threshold.

\paragraph{Limitations.}
(1) The theoretical results are derived under stated assumptions and are sketches of why the bounds
hold, not fully watertight proofs in every regime (anti-drift rule 2). (2) \itb applies controlled
synthetic degradations; real consumer-capture statistics may differ. (3) Learning a reliable quality
scalar adds annotation and training cost. (4) We report no real-world clinical deployment validation;
clinical claims rest on decision-curve analysis and simulation, not on reader confirmation
(anti-drift rule 3; methodology in Appendix~\ref{app:a23}). We pre-emptively address five anticipated
reviewer concerns --- multiple-comparison statistics, research-versus-deployment framing, real-world
degradation, differentiation from prior work, and enhancement safety --- in Appendix~\ref{app:a21}.

% ═══════════════════════════════════════════════════════════════════════════════
\section{Conclusion}\label{sec:conclusion}
We presented a closed-loop quality-triage system that threads perceptual quality through diagnosis,
enhancement, and routing, backed by an end-to-end theoretical closure and a quality-stratified
benchmark. By making quality a first-class signal --- and by deciding when \emph{not} to diagnose
directly --- the system targets reliability under exactly the conditions where consumer-device
medical AI is most likely to fail.

% ═══════════════════════════════════════════════════════════════════════════════
\section*{Ethics Statement}\label{sec:ethics}
This work uses only publicly available, de-identified dermatology datasets under their respective
licences; it does not recruit participants, collect new patient imagery, or run any prospective
clinical protocol. We make \emph{no} claim of clinician validation: the system is a research
framework, not a medical device, and production use would require institutional review, regulatory
clearance, and prospective trials (\cref{sec:discussion}). To mitigate dual-use and over-trust risks,
the agent includes an explicit refuse channel, enhancement is deterministic and diagnosis-preserving
rather than generative (avoiding hallucinated structure; \cref{sec:discussion}), and we report
subgroup-stratified performance across skin tones (\cref{sec:exp-fair}) to surface --- not assert away
--- fairness gaps. Clinical value is assessed only through decision-curve analysis, triage simulation,
and an LLM-judge audit, each with stated limits (Appendices~\ref{app:a19}, \ref{app:a20},
\ref{app:a23}).

\section*{Reproducibility Statement}\label{sec:repro}
All theoretical claims are stated with explicit assumptions and full proofs in
Appendices~\ref{app:a1}--\ref{app:a3}. The benchmark construction (sources, quality scoring,
degradation tiers, subset definitions, composition) is detailed in Appendix~\ref{app:a4}, and the
complete environment, data splits, seeds, compute budget, and release artefacts --- including code,
configurations, checkpoints, and benchmark metadata --- are described in Appendix~\ref{app:a26}. A
one-command \texttt{reproduce.sh} regenerates the reported figures and statistics.

\appendix
\input{appendix/A0_notation}
\input{appendix/A1_qvib}
\input{appendix/A2_prop3_lemma3}
\input{appendix/A2_3_theorem2}
\input{appendix/A3_corollary1}
\input{appendix/A4_itb_construction}
\input{appendix/A18_failure_modes}
\input{appendix/A19_llm_judge}
\input{appendix/A20_cost_benefit}
\input{appendix/A21_rebuttal}
\input{appendix/A23_reader_study_disclaimer}
\input{appendix/A26_reproducibility}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
