% ============================================================
% §6  Experimental Setup  (S version, 9-page short).
% Compressed from drafts/s6_setup.tex (716 w -> <=300 w).
% Drift contract: serves C1 (diagnosis-preserving enhancement) lever; this
% section is the evaluation substrate, NOT a contribution (benchmark demoted).
% Full benchmark construction -> Appendix A4.
% R-rules kept: 6 enhancement baselines, diffusion forbidden (R8), bootstrap
% 2000-sample 95% CI (R6), BMVC cited not reused (R10), anonymise (R4).
% Numbers copied verbatim from drafts/s6_setup.tex; none recomputed.
% ============================================================

\section{Experimental Setup}
\label{sec:setup}

\paragraph{Datasets.}
We evaluate on dermoscopic skin-lesion imagery from the public ISIC collection
(ISIC-2020), with a fixed train/validation/test split. Paired diagnostic
evaluations use the held-out paired subset ($n=3627$, with $117$
melanoma-positive cases; the full test split has $6{,}626$ images). The
diagnostic backbone is a frozen EfficientNet-B3 classifier (B3), held fixed
across all evaluations so that any change in diagnostic AUC is attributable to
the input rather than to retraining. All imagery is public; we run no reader study.

\paragraph{Degradation protocol.}
To make quality a known variable, we apply controlled degradations along five
physically parameterised dimensions: blur, brightness, completeness, colour
shift, and contrast. The main paired-evaluation protocol applies a
\emph{moderate} tier; the per-dimension $\times$ per-severity decision surface of
\S\ref{sec:decision-surface} sweeps each dimension across five severity levels.
Full parameters and the severity grid are deferred to Appendix~\ref{app:a4}.

\paragraph{Metrics.}
We report restoration fidelity as per-image PSNR and SSIM, and diagnostic
behaviour as the frozen-B3 AUC, the enhancement-induced $\Delta\mathrm{AUC}$, the
enhanced-vs-reference agreement rate, and the feature-level KL divergence
controlled by the DP-Loss. Every PSNR, AUC, $\Delta\mathrm{AUC}$, and agreement
figure carries a bootstrap $2000$-sample $95\%$ confidence interval; paired
comparisons add a McNemar test, and latency is reported in milliseconds per
image. A companion calibration study reports a complementary per-dimension
view, which we cite \citep{ANON-BMVC} but do not reuse.

\paragraph{Enhancement baselines.}
We compare against six publicly available deterministic restoration backbones
spanning convolutional and transformer designs: Restormer~\citep{zamir2022restormer},
NAFNet~\citep{chen2022simple}, MIRNet-v2~\citep{zamir2022mirnetv2},
SwinIR~\citep{liang2021swinir}, Uformer-B~\citep{wang2022uformer}, and
Real-ESRGAN~\citep{wang2021realesrgan}, each run zero-shot from official weights
under the identical protocol and frozen B3. We deliberately exclude diffusion and
other generative restorers: on dermoscopy they can hallucinate diagnostically
salient structure (invented pigment networks or vessels), a medical-safety hazard
we analyse in \S\ref{sec:discussion} rather than admit as a baseline.

